---
title: "Alluvial_plot_wetland_methanogens"
author: "Liupf"
date: "3/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Methanogen by top three familes across all sites with month (alluvial plot)


```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```
#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
library("factoextra")
library("FactoMineR") # PCA function
library("gplots") #heatmap
library("vegan") # prcomp
library('vegetarian') # normalize.rows function
library(dplyr)
library(tibble)
library(dendsort) #library(dendsort)
library(pheatmap)
```
#family level alluvial plot, top three methanogenes family

```{r}
library(ggalluvial)
library(reshape2)
library(dplyr)
library(alluvial)
```

```
# mapping file
map_file_clean_Cores.txt
# ASV file, 
feature_table_w_tax_clean_Cores_ge10.txt
# in bash
sed -e '1s/\./-/g' feature_table_w_tax_clean_Cores_ge10.txt >feature_table_w_tax_clean_Cores_ge10f.txt

# tax file
filtered_ge10_trimmed_tax_tab.txt
```
# read file
```{r}
map_file<-read.delim("map_file_clean_Cores.txt",header=T,row.names=1)
otus<-read.delim("feature_table_w_tax_clean_Cores_ge10f.txt",header=T,row.names=1, check.names=FALSE) 
taxa<-read.delim("filtered_ge10_trimmed_tax_tab.txt",header = T,row.names = 1) #tax_info, 
```

# add eco_dep column in the mapping file
```{r}
library(tidyr)
map_file$ecosite1=gsub("1|2|3","",map_file$ecosite)
map_file = unite(map_file, eco_dep, c(ecosite1, depth), remove=FALSE)
map_file = map_file[order(map_file$eco_dep), ,drop=FALSE] #M (D1-D6), N, OW, T is the order
```

#separate Bacteria and archaea
# data selction, separate Archea and Bacteria
```{r}
otus_Ba <- otus[grep("Bacteria", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ]
otus_Ar <- otus[grep("Archaea", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#map_file <- map_file[grep("Aug-|Aug_", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
#otus <- otus[, grep("Aug-|Aug_|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```


#claculate relative abundances
```{r}
library(gtools)
#bacteria
otus_Ba_ReAb=sweep(otus_Ba[,1:(ncol(otus_Ba)-1)], 2, colSums(otus_Ba[,1:(ncol(otus_Ba)-1)]), FUN="/")
otus_Ba_ReAb = otus_Ba_ReAb*100
otus_Ba_ReAb$taxonomy= otus_Ba$taxonomy
otus_Ba_ReAb=transform(merge(otus_Ba_ReAb,taxa, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
name=c(colnames(otus_Ar), colnames(taxa))

otus_Ba_ReAb=na.omit(otus_Ba_ReAb)
otus_Ba_ReAb$Genus <- sub("^$", "g_", otus_Ba_ReAb$Genus)
otus_Ba_ReAb$Family <- sub("^$", "f_", otus_Ba_ReAb$Family)
otus_Ba_ReAb$Order <- sub("^$", "o_", otus_Ba_ReAb$Order)
otus_Ba_ReAb$Class <- sub("^$", "c_", otus_Ba_ReAb$Class)
colnames(otus_Ba_ReAb)<-name
```

#archaea
```{r}
otus_Ar_ReAb=sweep(otus_Ar[,1:(ncol(otus_Ar)-1)], 2, colSums(otus_Ar[,1:(ncol(otus_Ar)-1)]), FUN="/")
otus_Ar_ReAb = otus_Ar_ReAb*100
otus_Ar_ReAb$taxonomy= otus_Ar$taxonomy
otus_Ar_ReAb=transform(merge(otus_Ar_ReAb,taxa, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
name=c(colnames(otus_Ar), colnames(taxa))

otus_Ar_ReAb=na.omit(otus_Ar_ReAb)
otus_Ar_ReAb$Genus <- sub("^$", "g_", otus_Ar_ReAb$Genus)
otus_Ar_ReAb$Family <- sub("^$", "f_", otus_Ar_ReAb$Family)
otus_Ar_ReAb$Order <- sub("^$", "o_", otus_Ar_ReAb$Order)
otus_Ar_ReAb$Class <- sub("^$", "c_", otus_Ar_ReAb$Class)
colnames(otus_Ar_ReAb)<-name



#
#feature_table_Arc_w_tax= merge(otus, Arc, by=0, all=FALSE)
#otus_D_ReAb = otus_D_ReAb [apply(otus_D_ReAb, 1, max)>=0.001,]
#otus_D_ReAb 
#write.table(data.frame("Sample_ID"=rownames(map_file_D),map_file_D),"map_file_Aug_D1_Cores.txt", sep="\t", row.names=FALSE, quote = FALSE)
```

#methanogens selection
#go to silva to check we use the same "name" for searching
https://www.arb-silva.de/search/

```{r}
#methanogens=c("Methanomicrobia","Methanobacteria","Methanococci","Methanopyri","Methanomassiliicoccales","Methanofastidiosales","WSA2","Methanofastidiosa","Verstraetearchaeia","Verstraetearchaeota","Methanomethyliales","Methanohydrogenales","Bathyarchaeia","Bathyarchaeota","Methanonatronarchaeia","Methanoflorentaceae","Methanoflorens","Syntrophoarchaeum","Rice Cluster II")

#canonical methanogens
#Cmethanogens=c("Methanomicrobia","Methanobacteria","Methanococci","Methanopyri","Methanomassiliicoccales","Methanofastidiosales","Methanofastidiosa","Verstraetearchaeia","Verstraetearchaeota","Methanomethyliales","Methanohydrogenales","Methanonatronarchaeia")

otus_methanogens = otus_Ar_ReAb[grep("Methanomicrobia|Methanobacteria|Methanococci|Methanopyri|Methanomassiliicoccales|Methanofastidiosales|WSA2|Methanonatronarchaeia|Methanoflorentaceae|Methanoflorens|Methanofastidiosa|Verstraetearchaeia|Verstraetearchaeota|Methanomethyliales|Methanohydrogenales|Bathyarchaeia|Bathyarchaeota|Syntrophoarchaeum|Rice Cluster II", otus_Ar_ReAb$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#Methanoflorens, Rice Cluster II

Cotus_methanogens = otus_Ar_ReAb[grep("Methanomicrobia|Methanobacteria|Methanococci|Methanopyri|Methanomassiliicoccales|Methanofastidiosales|Methanonatronarchaeia|Methanoflorentaceae|Methanoflorens|Methanofastidiosa|Verstraetearchaeia|Verstraetearchaeota|Methanomethyliales|Methanohydrogenales|Rice Cluster II", otus_Ar_ReAb$taxonomy, ignore.case=TRUE, invert=FALSE), ] 

```

#aggregate to family level
```{r}

#tax_tem_ar=otus_Ar_ReAb[,631:638]
Cotus_methanogens = unite(Cotus_methanogens, Ord_Fam, c(Order, Family), remove=FALSE)
Cotus_methanogens_Fam=transform(aggregate(x=Cotus_methanogens[,1:(ncol(Cotus_methanogens)-9)],by=list(Cotus_methanogens$Ord_Fam),FUN='sum'), row.names=Group.1,Group.1=NULL)
name=c(colnames(Cotus_methanogens[,1:(ncol(Cotus_methanogens)-9)]))
colnames(Cotus_methanogens_Fam)<-name
```

#choose the top4 family
```{r}
Cotus_methanogens_Fam$sum=apply(Cotus_methanogens_Fam[,1:630], 1, sum)
Cotus_methanogens_Fam$max=apply(Cotus_methanogens_Fam[,1:630], 1, max)
#order by sum, sort by sum is used since it is more clear
Cotus_methanogens_Fam=Cotus_methanogens_Fam[order(Cotus_methanogens_Fam$sum, decreasing = TRUE),]
Cotus_methanogens_Fam$Ord_Fam=rownames(Cotus_methanogens_Fam)
Cotus_methanogens_Fam=separate(Cotus_methanogens_Fam,Ord_Fam,sep = "_",into = c("Order", "Family"))
Cotus_methanogens_Fam_top4=Cotus_methanogens_Fam[1:4,]
```

#transform, make OTUs as the cloumn and samples as rows,
```{r}
Cotus_methanogens_Fam_top4=as.data.frame(t(Cotus_methanogens_Fam_top4[,1:630]))
Cotus_methanogens_Fam_top4=transform(merge(Cotus_methanogens_Fam_top4,map_file, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
```

#separate data by sites

#M1
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("M1", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#M2
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("M2", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#M3
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("M3", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#N1
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("N1", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#N2
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("N2", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#N3
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("N3", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#OW1
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("OW1", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#OW2
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("OW2", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#OW3
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("OW3", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```

#T1
```{r}
Fam_site=Cotus_methanogens_Fam_top4[grep("T1", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE),]
Fam_site=Fam_site[,c(1,2,3,4,8,10,12)]
```


#total
```{r}
Cotus_methanogens_Fam_top4=Cotus_methanogens_Fam_top4[,c(1,2,3,4,8,9,10,12)]
#M1_level5_archM<-melt(Cotus_methanogens_Fam_top4,id.vars=c("month","depth","core", "ecosite"))

M1_level5_archM=gather(Cotus_methanogens_Fam_top4, key="variable", value = "value",  -month, -depth, -core, -ecosite, factor_key= TRUE)

summarized_arch_for_alluvial<- group_by(M1_level5_archM, month, ecosite, depth, variable) %>% summarise(mean=mean(value))

summarized_arch_for_alluvial=summarized_arch_for_alluvial[grep("D1|D3|D6", summarized_arch_for_alluvial$depth,ignore.case=TRUE, invert=FALSE),]

#summarized_arch_for_alluvial =summarized_arch_for_alluvial [summarized_arch_for_alluvial$depth==c("D1","D3","D6"), ]

summarized_arch_for_alluvial$month<-factor(summarized_arch_for_alluvial$month,levels=c("May","June","July","Aug","Sept"))
```

```{r}
p=ggplot(data = summarized_arch_for_alluvial, 
       aes(x = month, y = mean, alluvium = variable)) +
  geom_alluvium(aes(fill = variable, colour = variable),alpha = .75, decreasing = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_grid(depth~ecosite, scales = "fixed") +theme(legend.position="bottom")+theme(legend.text = element_text( size = 12))+ ggtitle("Ecosites-methanogens dynamics") #change title

p
```
```{r}
ggsave(p,filename = "Alluvial_plot_ALLsites_methanogens.pdf", units = "in",height = 10, width = 16)
```



#y=time, xl=depth (xs=relative abundance)
#from Adrienne

```{r}
M1_level5_archM<-melt(Fam_site,id.vars=c("month","depth","core"))

summarized_arch_for_alluvial<-as.data.frame(group_by(M1_level5_archM,variable,month,depth)%>%summarize(mean=mean(value)))

summarized_arch_for_alluvial$month<-factor(summarized_arch_for_alluvial$month,levels=c("May","June","July","Aug","Sept"))

```
#choose subset for plotting depends on data completeness
```{r}
#M1, M3, N1, N2, N3, OW1, OW2, OW3, T1
summarized_arch_for_alluvial=summarized_arch_for_alluvial[summarized_arch_for_alluvial$depth==c("D1","D3","D6"), ] #1,3,6 dataset is full for all month
```

```{r}
#M2, 
summarized_arch_for_alluvial=summarized_arch_for_alluvial[summarized_arch_for_alluvial$depth==c("D1","D3","D5"), ] #1,3,6 dataset is full for all month
```


# try the ggalluvial
#https://cran.r-project.org/web/packages/ggalluvial/vignettes/ggalluvial.html
```{r}
p=ggplot(data = summarized_arch_for_alluvial, 
       aes(x = month, y = mean, alluvium = variable)) +
  geom_alluvium(aes(fill = variable, colour = variable),alpha = .75, decreasing = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~ depth, scales = "fixed",3) +
  ggtitle("Ecosite-T1-methanogens dynamics") #change title
#M1, M2, M3, N1, N2,N3, OW1, OW2
#scale_x_continuous(breaks = seq(2003, 2013, 2)) +
```
```{r}
ggsave(p,filename = "Alluvial_plot_T1_methanogens.pdf", units = "in",height = 10, width = 10)
```

```{r}
ggplot(summarized_methanogens_for_alluvial_just136,aes(x=timepoint,y=mean,alluvium=taxon_short))+geom_alluvium(aes(fill = taxon_short, colour = taxon_short),alpha = .75, decreasing = FALSE)+scale_fill_manual(values=colors)+scale_color_manual(values=colors)+facet_grid(depth~.)+xlab(“”)+ylab(“mean relative abundance”)+theme_bw()+theme(legend.title = element_blank())+geom_flow()
```

#ridgeline plot, joyplot
x for time, y for relative abundance, Depth by each pane/layer, color for taxa,

#when we got this, try to smaller ASV level, with the three top family, alluvial plot
separate by sites
x, month, y relative abundance, stack depth
```{r}

```




#pheatmap plotting on family level


#
```{r}
#based on old alluvial package, not nice

arch_D1<-subset(summarized_arch_for_alluvial,depth=="D1")[,c(1,2,4)]
arch_D2<-subset(summarized_arch_for_alluvial,depth=="D2")[,c(1,2,4)]
arch_D3<-subset(summarized_arch_for_alluvial,depth=="D3")[,c(1,2,4)]
arch_D4<-subset(summarized_arch_for_alluvial,depth=="D4")[,c(1,2,4)]
arch_D5<-subset(summarized_arch_for_alluvial,depth=="D5")[,c(1,2,4)]
arch_D6<-subset(summarized_arch_for_alluvial,depth=="D6")[,c(1,2,4)]
alluvial_ts(arch_D1, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D2, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D3, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D4, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D5, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D6, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)

```
