---
title: "Alluvial_plot_wetland_methanotroph"
author: "Liupf"
date: "3/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Methanogen by top three familes across all sites with month (alluvial plot)

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#purpose
#Methanogen (row) by OTU across samples (columns)- sum the total and sort, so we can see most abundant bugs, fine do archaea. Can organize by months if too big to show

```{r}

```
```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```
#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
library("factoextra")
library("FactoMineR") # PCA function
library("gplots") #heatmap
library("vegan") # prcomp
library('vegetarian') # normalize.rows function
library(dplyr)
library(tidyr)
library(tibble)
library(dendsort) #library(dendsort)
library(pheatmap)
library(ggalluvial)
```

#color code setup
```{r}
library(RColorBrewer)
bluepal<-brewer.pal(6,"Blues") #open water
orangepal<-brewer.pal(6,"Oranges") # mud
greenpal1<-brewer.pal(6,"Greens") # Typha
greenpal2<-brewer.pal(6,"YlGn") # Nelumbo
pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[2], greenpal1[3], greenpal1[4], bluepal[4], bluepal[5], bluepal[6], greenpal2[6])


```

```

```
# read file
```{r}
wd<-read.delim("OWC2018sampling_water_depth_liupf1.txt",header=T,sep = "\t")
 
```

#total
```{}
Cotus_methanotroph_Fam_top3=Cotus_methanotroph_Fam_top3[,c(1,2,3,7,8,9,11)]
#M1_level5_archM<-melt(Cotus_methanotroph_top3,id.vars=c("month","depth","core", "ecosite"))

M1_level5_archM=gather(Cotus_methanotroph_Fam_top3, key="variable", value = "value",  -month, -depth, -core, -ecosite, factor_key= TRUE)

summarized_arch_for_alluvial<- group_by(M1_level5_archM, month, ecosite, depth, variable) %>% summarise(mean=mean(value))

summarized_arch_for_alluvial=summarized_arch_for_alluvial[grep("D1|D3|D6", summarized_arch_for_alluvial$depth,ignore.case=TRUE, invert=FALSE),]

#summarized_arch_for_alluvial =summarized_arch_for_alluvial [summarized_arch_for_alluvial$depth==c("D1","D3","D6"), ]

summarized_arch_for_alluvial$month<-factor(summarized_arch_for_alluvial$month,levels=c("May","June","July","Aug","Sept"))
```

```{r}
wd$month<-factor(wd$month,levels=c("May","June","July","Aug","Sept"))
wd$variable="depth"
```

```{r}
p=ggplot(data = wd,  aes(x = month, y = W_depth_cm, alluvium = variable)) +
  geom_alluvium(aes(fill = variable, colour = variable),alpha = .75, decreasing = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~Site,ncol=10) +theme(legend.position="bottom")+
  ggtitle("Ecosites-covering water dynamics") #change title

p

```

```{r}

wd$month<-factor(wd$month,levels=c("May","June","July","Aug","Sept"))
wd$variable="depth"
```

```{r}
p1=ggplot(data = wd,  aes(x = month, y = W_depth_cm, alluvium = variable)) +
  geom_alluvium(aes(fill = variable, colour = variable),alpha = .75, decreasing = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_manual(values=NA) +scale_color_manual(values=c("gray30"))+
  facet_wrap(~Site,ncol=10) +theme(legend.position="none")+
  ggtitle("Ecosites-covering water dynamics") #change title


p1
```



#save
```{r}
ggsave(p,filename = "Alluvial_plot_ALLsites_methanogens.pdf", units = "in",height = 10, width = 16)
```



#y=time, xl=depth (xs=relative abundance)
#from Adrienne

```{r}
M1_level5_archM=gather(Fam_site, key="variable", value = "value",  -month, -depth, -core, -ecosite, factor_key= TRUE)

summarized_arch_for_alluvial<- group_by(M1_level5_archM, month, ecosite, depth, variable) %>% summarise(mean=mean(value))

summarized_arch_for_alluvial=summarized_arch_for_alluvial[grep("D1|D3|D6", summarized_arch_for_alluvial$depth,ignore.case=TRUE, invert=FALSE),]

#summarized_arch_for_alluvial =summarized_arch_for_alluvial [summarized_arch_for_alluvial$depth==c("D1","D3","D6"), ]

summarized_arch_for_alluvial$month<-factor(summarized_arch_for_alluvial$month,levels=c("May","June","July","Aug","Sept"))
```
#choose subset for plotting depends on data completeness
```{r}
#M1, M3, N1, N2, N3, OW1, OW2, OW3, T1
#summarized_arch_for_alluvial=summarized_arch_for_alluvial[summarized_arch_for_alluvial$depth==c("D1","D3","D6"), ] #1,3,6 dataset is full for all month
```

```{r}
#M2, 
#summarized_arch_for_alluvial=summarized_arch_for_alluvial[summarized_arch_for_alluvial$depth==c("D1","D3","D5"), ] #1,3,6 dataset is full for all month
```


# try the ggalluvial
#https://cran.r-project.org/web/packages/ggalluvial/vignettes/ggalluvial.html
```{r}
p=ggplot(data = summarized_arch_for_alluvial, 
       aes(x = month, y = mean, alluvium = variable)) +
  geom_alluvium(aes(fill = variable, colour = variable),alpha = .75, decreasing = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~ depth, scales = "fixed",3) +
    ggtitle("Ecosite-T1-methanotroph dynamics") #change title
#M1, M2, M3, N1, N2,N3, OW1, OW2
#scale_x_continuous(breaks = seq(2003, 2013, 2)) +
p
```
```{r}
ggsave(p,filename = "Alluvial_plot_T1_Methanotroph.pdf", units = "in",height = 10, width = 10)
```

```{r}
ggplot(summarized_methanogens_for_alluvial_just136,aes(x=timepoint,y=mean,alluvium=taxon_short))+geom_alluvium(aes(fill = taxon_short, colour = taxon_short),alpha = .75, decreasing = FALSE)+scale_fill_manual(values=colors)+scale_color_manual(values=colors)+facet_grid(depth~.)+xlab(“”)+ylab(“mean relative abundance”)+theme_bw()+theme(legend.title = element_blank())+geom_flow()
```

#ridgeline plot, joyplot
x for time, y for relative abundance, Depth by each pane/layer, color for taxa,

#when we got this, try to smaller ASV level, with the three top family, alluvial plot
separate by sites
x, month, y relative abundance, stack depth
```{r}

```




#pheatmap plotting on family level


#
```{r}
#based on old alluvial package, not nice

arch_D1<-subset(summarized_arch_for_alluvial,depth=="D1")[,c(1,2,4)]
arch_D2<-subset(summarized_arch_for_alluvial,depth=="D2")[,c(1,2,4)]
arch_D3<-subset(summarized_arch_for_alluvial,depth=="D3")[,c(1,2,4)]
arch_D4<-subset(summarized_arch_for_alluvial,depth=="D4")[,c(1,2,4)]
arch_D5<-subset(summarized_arch_for_alluvial,depth=="D5")[,c(1,2,4)]
arch_D6<-subset(summarized_arch_for_alluvial,depth=="D6")[,c(1,2,4)]
alluvial_ts(arch_D1, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D2, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D3, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D4, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D5, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)
alluvial_ts(arch_D6, wave = 1.0, ygap = 4, lwd = 1, alpha = 0.7, rankup = FALSE, plotdir = "up", grid=TRUE, border = "dark grey", xmargin = 01.0, lab.cex = 0.8)

```
