#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
```

```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```

# mapping file
map_file_clean_Cores.txt
# ASV file, 
feature_table_w_tax_clean_Cores_ge10.txt
# in bash
sed -e '1s/\./-/g' feature_table_w_tax_clean_Cores_ge10.txt >feature_table_w_tax_clean_Cores_ge10f.txt

# tax file
filtered_ge10_trimmed_tax_tab.txt

# read file
```{r}
map_file<-read.delim("map_file_clean_Cores.txt",header=T,row.names=1)

otus<-read.delim("feature_table_w_tax_clean_Cores_ge10f.txt",header=T,row.names=1, check.names=FALSE) 

taxa<-read.delim("filtered_ge10_trimmed_tax_tab.txt",header = T,row.names = 1) #tax_info, 

```

# add eco_dep column in the mapping file
```{r}
library(tidyr)
map_file$ecosite1=gsub("1|2|3","",map_file$ecosite)
map_file = unite(map_file, eco_dep, c(ecosite1, depth), remove=FALSE)
map_file = map_file[order(map_file$eco_dep), ,drop=FALSE] #M (D1-D6), N, OW, T is the order
```

#color code setup
```{r}
library(RColorBrewer)
bluepal<-brewer.pal(6,"Blues") #open water
orangepal<-brewer.pal(6,"Oranges") # mud
greenpal1<-brewer.pal(6,"Greens") # Typha
greenpal2<-brewer.pal(6,"YlGn") # Nelumbo
pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[2], greenpal1[3], greenpal1[4], bluepal[4], bluepal[5], bluepal[6], greenpal2[6])


```
# taxmat
```{r}
taxmat<-as.matrix(taxa)
TAX = tax_table(taxmat) #covert to phyoseq object, OTU table
```


# August data selction
```{r}
map_file <- map_file[grep("Aug-|Aug_", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus <- otus[, grep("Aug-|Aug_|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# August-D1 selction, move to  phyloseq section after data selection
```{r}
map_file_D <- map_file[grep("-D1-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D1-|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```
##data for stamps
#export metafile and otu file
```{r}
library(gtools)
otus_D_ReAb=sweep(otus_D[,1:(ncol(otus_D)-1)], 2, colSums(otus_D[,1:(ncol(otus_D)-1)]), FUN="/")
otus_D_ReAb = otus_D_ReAb [apply(otus_D_ReAb, 1, max)>=0.001,]
otus_D_ReAb 

write.table(data.frame("Sample_ID"=rownames(map_file_D),map_file_D),"map_file_Aug_D1_Cores.txt", sep="\t", row.names=FALSE, quote = FALSE)
write.table(data.frame("OTU_ID"=rownames(otus_D_ReAb ), otus_D_ReAb ),"feature_table_Aug_D1_Cores_ge10.txt", sep="\t", row.names=FALSE, quote = FALSE)
```

# August-D2 selction, move to  phyloseq section after data selection
```{r}
map_file_D <- map_file[grep("-D2-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D2-", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# August-D3 selction, move to  phyloseq section after data selection
```{r}
map_file_D <- map_file[grep("-D3-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D3-", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# August-D4 selction
```{r}
map_file_D <- map_file[grep("-D4-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D4-", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# August-D5 selction
```{r}
map_file_D <- map_file[grep("-D5-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D5-", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# August-D6 selction, move to  phyloseq section after data selection
```{r}
map_file_D <- map_file[grep("-D6-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus[, grep("-D6-", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```
# move to  phyloseq section after data selection


# phyloseq section and plot
# 
# otu
```{r}
otumat<-as.matrix(otus_D[[,-ncol(otus_D)])
OTU = otu_table(otumat, taxa_are_rows = TRUE)
physeq<-phyloseq(OTU,TAX) 

sampledata<-sample_data(map_file_D)
mgd<-merge_phyloseq(physeq,sampledata)
```

#plot

```{r}
mgd.ord <- ordinate(mgd, "NMDS", "bray")
```

```{r}
theme_set(theme_bw())
pmds=plot_ordination(mgd, mgd.ord, color="ecosite", shape="depth") #
pmds=pmds + geom_point(size=1.5)+scale_color_manual(values=pal) #
pmds
```


