---
title: "OWC_methanotrophs_OTUs_relative abundance plot"
author: "Liupf"
date: "3/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#purpose
#Methanogen (row) by OTU across samples (columns)- sum the total and sort, so we can see most abundant bugs, fine do archaea. Can organize by months if too big to show

```{r}

```
```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```
#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
library("factoextra")
library("FactoMineR") # PCA function
library("gplots") #heatmap
library("vegan") # prcomp
library('vegetarian') # normalize.rows function
library(dplyr)
library(tidyr)
library(tibble)
library(dendsort) #library(dendsort)
library(pheatmap)
```

#color code setup
```{r}
library(RColorBrewer)
bluepal<-brewer.pal(6,"Blues") #open water
orangepal<-brewer.pal(6,"Oranges") # mud
greenpal1<-brewer.pal(6,"Greens") # Typha
greenpal2<-brewer.pal(6,"YlGn") # Nelumbo
pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[2], greenpal1[3], greenpal1[4], bluepal[4], bluepal[5], bluepal[6], greenpal2[6])


```

```
# mapping file
map_file_clean_Cores.txt
# ASV file, 
feature_table_w_tax_clean_Cores_ge10.txt
# in bash
sed -e '1s/\./-/g' feature_table_w_tax_clean_Cores_ge10.txt >feature_table_w_tax_clean_Cores_ge10f.txt

# tax file
filtered_ge10_trimmed_tax_tab.txt
```
# read file
```{r}
map_file<-read.delim("map_file_clean_Cores.txt",header=T,row.names=1)
otus<-read.delim("feature_table_w_tax_clean_Cores_ge10f.txt",header=T,row.names=1, check.names=FALSE) 
taxa<-read.delim("filtered_ge10_trimmed_tax_tab.txt",header = T,row.names = 1) #tax_info, 
```

# add eco_dep column in the mapping file
```{r}
library(tidyr)
map_file$ecosite1=gsub("1|2|3","",map_file$ecosite)
map_file = unite(map_file, eco_dep, c(ecosite1, depth), remove=FALSE)
map_file = map_file[order(map_file$eco_dep), ,drop=FALSE] #M (D1-D6), N, OW, T is the order
```

#separate Bacteria and archaea
# data selction, separate Archea and Bacteria
```{r}
otus_Ba <- otus[grep("Bacteria", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ]
otus_Ar <- otus[grep("Archaea", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#map_file <- map_file[grep("Aug-|Aug_", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
#otus <- otus[, grep("Aug-|Aug_|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```


#claculate relative abundances
```{r}
library(gtools)
#bacteria
otus_Ba_ReAb=sweep(otus_Ba[,1:(ncol(otus_Ba)-1)], 2, colSums(otus_Ba[,1:(ncol(otus_Ba)-1)]), FUN="/")
otus_Ba_ReAb = otus_Ba_ReAb*100
otus_Ba_ReAb$taxonomy= otus_Ba$taxonomy
otus_Ba_ReAb=transform(merge(otus_Ba_ReAb,taxa, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
name=c(colnames(otus_Ar), colnames(taxa))

otus_Ba_ReAb=na.omit(otus_Ba_ReAb)
otus_Ba_ReAb$Genus <- sub("^$", "g_", otus_Ba_ReAb$Genus)
otus_Ba_ReAb$Family <- sub("^$", "f_", otus_Ba_ReAb$Family)
otus_Ba_ReAb$Order <- sub("^$", "o_", otus_Ba_ReAb$Order)
otus_Ba_ReAb$Class <- sub("^$", "c_", otus_Ba_ReAb$Class)
colnames(otus_Ba_ReAb)<-name
```

#archaea
```{r}
otus_Ar_ReAb=sweep(otus_Ar[,1:(ncol(otus_Ar)-1)], 2, colSums(otus_Ar[,1:(ncol(otus_Ar)-1)]), FUN="/")
otus_Ar_ReAb = otus_Ar_ReAb*100
otus_Ar_ReAb$taxonomy= otus_Ar$taxonomy
otus_Ar_ReAb=transform(merge(otus_Ar_ReAb,taxa, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
name=c(colnames(otus_Ar), colnames(taxa))

otus_Ar_ReAb=na.omit(otus_Ar_ReAb)
otus_Ar_ReAb$Genus <- sub("^$", "g_", otus_Ar_ReAb$Genus)
otus_Ar_ReAb$Family <- sub("^$", "f_", otus_Ar_ReAb$Family)
otus_Ar_ReAb$Order <- sub("^$", "o_", otus_Ar_ReAb$Order)
otus_Ar_ReAb$Class <- sub("^$", "c_", otus_Ar_ReAb$Class)
colnames(otus_Ar_ReAb)<-name

#tax_tem_ar=otus_Ar_ReAb[,631:638]
#otus_Ar=transform(aggregate(x=otus_Ar[,1:(ncol(otus_Ar)-8)],by=list(otus_Ar$Family),FUN='sum'), row.names=Group.1,Group.1=NULL)
#colnames(otus_Ar)<-name

#
#feature_table_Arc_w_tax= merge(otus, Arc, by=0, all=FALSE)
#otus_D_ReAb = otus_D_ReAb [apply(otus_D_ReAb, 1, max)>=0.001,]
#otus_D_ReAb 
#write.table(data.frame("Sample_ID"=rownames(map_file_D),map_file_D),"map_file_Aug_D1_Cores.txt", sep="\t", row.names=FALSE, quote = FALSE)
```

```{r}
#done
#write.table(data.frame("OTU_ID"=rownames(otus_Ar_ReAb ), otus_Ar_ReAb ),"ASV_Cores_ge10_Archaea_Reab.txt", sep="\t", row.names=FALSE, quote = FALSE)
#write.table(data.frame("OTU_ID"=rownames(otus_Ba_ReAb ), otus_Ba_ReAb ),"ASV_Cores_ge10_Bacteria_Reab.txt", sep="\t", row.names=FALSE, quote = FALSE)
```


#methantroph selection
#go to silva to check we use the same "name" for searching
https://www.arb-silva.de/search/

```{r}
#methanortroph from Bacteria

#"Methylomonaceae|Methylococcaceae|Methylothermaceae|Crenothrix|Clonothrix|Methylocystaceae|Methyloligellaceae|Methylopilaceae|Methylobacterium|Methylocapsa|Methylocella|Methylocystis|Methyloferula|Methylorosula|Methylosinus|Methylovirgula|Methylacidophilales|Methylacidophilaceae|Methylomirabilis|Methylacidimicrobium|Methylospira"


#Methylococcales (Methylomonaceae|Methylococcaceae|Methylothermaceae|Crenothrix)
#not included:USCα|USCγ, NC-10, 
#Beijerinckiaceae (use Methylocella,Methylobacterium|Methylocapsa|Methylocella|Methylocystis|Methyloferula|Methylorosula|Methylosinus|Methylovirgula)

#typical methanortroph from Bacteria
Cotus_methanotroph = otus_Ba_ReAb[grep("Methylomonaceae|Methylococcaceae|Methylothermaceae|Crenothrix|Clonothrix|Methylocystaceae|Methyloligellaceae|Methylopilaceae|Methylobacterium|Methylocapsa|Methylocella|Methylocystis|Methyloferula|Methylorosula|Methylosinus|Methylovirgula|Methylacidophilales|Methylacidophilaceae|Methylomirabilis|Methylacidimicrobium|Methylospira", otus_Ba_ReAb$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#Methylococcales
```

#sort row order
```{r}
#feature_table_Arc_w_tax= merge(otus, Arc, by=0, all=FALSE)
#get the sum of each ASV across ALL sample
Cotus_methanotroph_f= Cotus_methanotroph [apply(Cotus_methanotroph [,1:630], 1, max)>=0.1,] #0.1% 
Cotus_methanotroph_f$sum=apply(Cotus_methanotroph_f[,1:630], 1, sum)
Cotus_methanotroph_f$max=apply(Cotus_methanotroph_f[,1:630], 1, max)
#order by sum
#Cotus_methanotroph_f=Cotus_methanotroph_f[order(Cotus_methanotroph_f$sum, decreasing = TRUE),]
#sort by max
#Cotus_methanotroph_f=Cotus_methanotroph_f[order(Cotus_methanotroph_f$max, decreasing = TRUE),]
#Cotus_methanotroph_f_tem=Cotus_methanotroph_f[,631:640]
Cotus_methanotroph_f = unite(Cotus_methanotroph_f, Ord_Fam, c(Order, Family), remove=FALSE)
```
# All-Archaea-by month
#May
```{r}
map_file_D <- map_file[grep("May", row.names(map_file), ignore.case=TRUE, invert=FALSE),]

otus_D <- Cotus_methanotroph_f[, grep("May|taxonomy|Ord_Fam|Order|Family|Genus", colnames(Cotus_methanotroph_f), ignore.case=TRUE, invert=FALSE)]
otus_D= otus_D[apply(otus_D[ ,1:(ncol(otus_D)-5)], 1, max)>=0.01,] #0.1% is mess
otus_D$sum=apply(otus_D[,1:(ncol(otus_D)-5)], 1, sum)
otus_D$max=apply(otus_D[,1:(ncol(otus_D)-6)], 1, max)

#order by sum, sort by sum is used since it is more clear
otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$Family, otus_D$Genus, otus_D$sum, decreasing = TRUE),]

#sort by max
#otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$max, decreasing = TRUE),]
```
#move to pheatmap ploting after each data prep of month

# June
```{r}
map_file_D <- map_file[grep("June", row.names(map_file), ignore.case=TRUE, invert=FALSE),]

otus_D <- Cotus_methanotroph_f[, grep("June|taxonomy|Ord_Fam|Order|Family|Genus", colnames(Cotus_methanotroph_f), ignore.case=TRUE, invert=FALSE)]
otus_D= otus_D[apply(otus_D[ ,1:(ncol(otus_D)-5)], 1, max)>=0.01,] #0.1% is mess
otus_D$sum=apply(otus_D[,1:(ncol(otus_D)-5)], 1, sum)
otus_D$max=apply(otus_D[,1:(ncol(otus_D)-6)], 1, max)

#order by sum, sort by sum is used since it is more clear
otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$Family, otus_D$Genus, otus_D$sum, decreasing = TRUE),]

```

# July
```{r}
map_file_D <- map_file[grep("July", row.names(map_file), ignore.case=TRUE, invert=FALSE),]

otus_D <- Cotus_methanotroph_f[, grep("July|taxonomy|Ord_Fam|Order|Family|Genus", colnames(Cotus_methanotroph_f), ignore.case=TRUE, invert=FALSE)]
otus_D= otus_D[apply(otus_D[ ,1:(ncol(otus_D)-5)], 1, max)>=0.01,] #0.1% is mess
otus_D$sum=apply(otus_D[,1:(ncol(otus_D)-5)], 1, sum)
otus_D$max=apply(otus_D[,1:(ncol(otus_D)-6)], 1, max)

#order by sum, sort by sum is used since it is more clear
otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$Family, otus_D$Genus, otus_D$sum, decreasing = TRUE),]
```


# Aug
```{r}
map_file_D <- map_file[grep("Aug", row.names(map_file), ignore.case=TRUE, invert=FALSE),]

otus_D <- Cotus_methanotroph_f[, grep("Aug|taxonomy|Ord_Fam|Order|Family|Genus", colnames(Cotus_methanotroph_f), ignore.case=TRUE, invert=FALSE)]
otus_D= otus_D[apply(otus_D[ ,1:(ncol(otus_D)-5)], 1, max)>=0.01,] #0.1% is mess
otus_D$sum=apply(otus_D[,1:(ncol(otus_D)-5)], 1, sum)
otus_D$max=apply(otus_D[,1:(ncol(otus_D)-6)], 1, max)

#order by sum, sort by sum is used since it is more clear
otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$Family, otus_D$Genus, otus_D$sum, decreasing = TRUE),]
```


# Sept
```{r}
map_file_D <- map_file[grep("Sept", row.names(map_file), ignore.case=TRUE, invert=FALSE),]

otus_D <- Cotus_methanotroph_f[, grep("Sept|taxonomy|Ord_Fam|Order|Family|Genus", colnames(Cotus_methanotroph_f), ignore.case=TRUE, invert=FALSE)]
otus_D= otus_D[apply(otus_D[ ,1:(ncol(otus_D)-5)], 1, max)>=0.01,] #0.1% is mess
otus_D$sum=apply(otus_D[,1:(ncol(otus_D)-5)], 1, sum)
otus_D$max=apply(otus_D[,1:(ncol(otus_D)-6)], 1, max)

#order by sum, sort by sum is used since it is more clear
otus_D=otus_D[order(otus_D$Ord_Fam, otus_D$Family, otus_D$Genus, otus_D$sum, decreasing = TRUE),]

```


#pheatmap plotting

#break setting
```{r}
df = as.matrix(otus_D[,1:(ncol(otus_D)-7)])
#quantile.range <- quantile(as.matrix(df), probs = seq(0, 1, 0.01))
#mat_breaks <- seq(quantile.range["5%"], quantile.range["95%"], 0.001)
#color  <- colorRampPalette(c("blue", "white","firebrick3"))(length(mat_breaks) - 1) #c("royalblue4","#FFFFBF","orangered2"))

#color = colorRampPalette(c("white","darkred"))(400) #,"gray", firebrick3
color = colorRampPalette(c("white","red"))(400) #,"gray", firebrick3

#no cluster for col, and row
#drows = dist(as.matrix(decostand(df, "hellinger")), method = "euclidean") #decostand(OTUtable_t.na, "hellinger")
#dcols = dist(t(as.matrix(decostand(df, "hellinger"))), method = "euclidean") #decostand(OTUtable_t.na, "hellinger")
#sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
#mat_cluster_cols <- hclust(dcols)
#mat_cluster_rows <- hclust(drows)

#get row annotation
my_otus_row1=subset(otus_D, select=Order) 
my_otus_row2=subset(otus_D, select=Family) 
annotation_row =data.frame(Order=my_otus_row2, Family=my_otus_row1, row.names = rownames(otus_D))
 
#get column annotation---->sample sites
#use match to reorder df1, DF1[match(rownames(DF2), rownames(DF1)), ]
map_file_D=map_file_D[match(colnames(otus_D[,1:(ncol(otus_D)-7)]),rownames(map_file_D)), ] 
my_sample_col= subset(map_file_D, select=ecosite)

#get column annotation---->Depth
#use match to reorder df1, DF1[match(rownames(DF2), rownames(DF1)), ]
my_sample_col2= subset(map_file_D, select=depth)

#merge annotation_col
annotation_col =data.frame(depth=my_sample_col2,ecosite=my_sample_col, row.names = rownames(map_file_D))

#pal
#brewer.pal(n = 6, name = "Dark2") for the taxon
#brewer.pal(n = 9, name = "Greys") for depth
#barplot(c(1:9),cool=brewer.pal(n = 9, name = "Greys"))

#annotation colors
anno_colors= list(ecosite = c(M1=orangepal[4], M2=orangepal[5], M3=orangepal[6], N1=greenpal1[4], N2=greenpal1[5], N3=greenpal1[6], OW1=bluepal[4], OW2=bluepal[5], OW3=bluepal[6], T1=greenpal2[2]),  depth=c(D1="#D9D9D9", D2="#BDBDBD" , D3="#969696", D4="#737373", D5="#525252", D6="#252525"),Family=c(Methylococcaceae="#1B9E77", Methylomonaceae="#D95F02", Methyloligellaceae="#E7298A",Beijerinckiaceae= "#7570B3"), Order=c(Rhizobiales="#E7298A",Methylococcales="#E6AB02")) #list specify the annotation names

#
#, Rhizobiales="#E7298A"
```


#pheatmap plot
```{r, echo = FALSE}
pheatmap(df,cluster_rows=FALSE,cluster_cols=FALSE,labels_row=otus_D$Genus, cellwidth=8,cellheight=8,fontsize = 6, color = color, annotation_row =annotation_row,annotation_col = annotation_col,annotation_colors = anno_colors)
```

#pheatmap tricks
```{r}
#df = as.matrix(otus_D[,1:(ncol(otus_D)-6)])
#pheatmap(df,cluster_rows=FALSE,cluster_cols=FALSE,labels_row=otus_D$Ord_Fam, cellwidth=8,cellheight=8,fontsize = 6)
#pheatmap(df,color = color,breaks= mat_breaks,cellwidth=8,cellheight=5, show_colnames=TRUE, fontsize=6)
#https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/
#pheatmap(df,cluster_rows=FALSE,labels_row=otus_D$Ord_Fam, cellwidth=8,cellheight=8,fontsize = 6, clustering_distance_cols=dcols, clustering_method="complete", cluster_cols= mat_cluster_cols)
#pheatmap(df,color = color,breaks= mat_breaks,cellwidth=8,cellheight=5, show_colnames=TRUE, fontsize=6)
#https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/
#https://stackoverflow.com/questions/41628450/r-pheatmap-change-annotation-colors-and-prevent-graphics-window-from-popping-up
```

#pheatmap(as.matrix(percentages),color = color,breaks= mat_breaks,clustering_distance_rows = drows, clustering_distance_cols=dcols, clustering_method="complete",cellwidth=8,cellheight=5, cluster_cols= mat_cluster_cols,cluster_rows = mat_cluster_rows, show_colnames=TRUE, fontsize=6)


#transform, make OTUs as the cloumn and samples as rows,
```{r}


```

#family level alluvial plot, top three methanogenes family
```{r}


```

# All-by_Depth-Archaea
#D1
```{r}
#map_file_D <- map_file[grep("-D1-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D1-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]

```
#move to pca selection and ploting after each data prep of one depth

# D-2
```{r}
#map_file_D <- map_file[grep("-D2-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D2-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]
```

# D-3
```{r}
#map_file_D <- map_file[grep("-D3-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D3-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]
```


# D-4
```{r}
#map_file_D <- map_file[grep("-D4-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D4-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]
```


# D-5
```{r}
#map_file_D <- map_file[grep("-D5-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D5-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]
```


# D-6
```{r}
#map_file_D <- map_file[grep("-D6-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- Cotus_methanogens[, grep("-D6-|taxonomy|Ord_Fam|Order|Family", colnames(Cotus_methanogens), ignore.case=TRUE, invert=FALSE)]
```


