#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
```

```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```

# mapping file
map_file_clean_Cores.txt
# ASV file, 
feature_table_w_tax_clean_Cores_ge10.txt
# in bash
sed -e '1s/\./-/g' feature_table_w_tax_clean_Cores_ge10.txt >feature_table_w_tax_clean_Cores_ge10f.txt

# tax file
filtered_ge10_trimmed_tax_tab.txt

# read file
```{r}
map_file<-read.delim("map_file_clean_Cores.txt",header=T,row.names=1)
otus<-read.delim("feature_table_w_tax_clean_Cores_ge10f.txt",header=T,row.names=1, check.names=FALSE) 
taxa<-read.delim("filtered_ge10_trimmed_tax_tab.txt",header = T,row.names = 1) #tax_info, 
taxa$Family <- sub("^$", "Unclassified", taxa$Family)
```

# add eco_dep column in the mapping file
```{r}
library(tidyr)
map_file$ecosite1=gsub("1|2|3","",map_file$ecosite)
map_file = unite(map_file, eco_dep, c(ecosite1, depth), remove=FALSE)
map_file = map_file[order(map_file$eco_dep), ,drop=FALSE] #M (D1-D6), N, OW, T is the order
```

#color code setup
```{r}
library(RColorBrewer)
bluepal<-brewer.pal(6,"Blues") #open water
orangepal<-brewer.pal(6,"Oranges") # mud
greenpal1<-brewer.pal(6,"Greens") # Typha
greenpal2<-brewer.pal(6,"YlGn") # Nelumbo
#pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[2], greenpal1[3], greenpal1[4], bluepal[4], bluepal[5], bluepal[6], greenpal2[6])


```

# separate all depth, keep all ecosite and month
# data selction, separate Archea and Bacteria
```{r}
#otus_Ar <- otus[grep("Archaea", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#map_file <- map_file[grep("Aug-|Aug_", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
#otus <- otus[, grep("Aug-|Aug_|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

#aggregate to Family level
```{r}
otus_Ba <- otus[grep("Bacteria", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ]
require(gtools) 
#name=c(colnames(otus_Ba),"Domain","Phylum","Class","Order","Family","Genus","Species")
name=c(colnames(otus_Ba[,-ncol(otus_Ba)]))
otus_Ba=transform(merge(otus_Ba,taxa, by='row.names',sort=F, all.x=TRUE),row.names=Row.names, Row.names=NULL)
#OTU_Tax16SREab=merge(OTU_16SREab,Taxo16S,by='row.names',sort=F)
#OTU_16SREab=sweep(OTU_16S[,1:54], 2, colSums(OTU_16S[,1:54]), FUN="/")
otus_Ba=na.omit(otus_Ba)
#otus_Ba$Family <- sub("^$", "Unclassified", otus_Ba$Family)
otus_Ba=transform(aggregate(x=otus_Ba[,1:(ncol(otus_Ba)-8)],by=list(otus_Ba$Family),FUN='sum'), row.names=Group.1,Group.1=NULL)
colnames(otus_Ba)<-name
```

# All-month-Bacterial
# D-1
```{r}
map_file_D <- map_file[grep("-D1-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D1-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```

#D-2
```{r}
map_file_D <- map_file[grep("-D2-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D2-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```
#D-3
```{r}
map_file_D <- map_file[grep("-D3-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D3-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```
#D-4
```{r}
map_file_D <- map_file[grep("-D4-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D4-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```
#D-5
```{r}
map_file_D <- map_file[grep("-D5-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D5-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```
#D-6
```{r}
map_file_D <- map_file[grep("-D6-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D6-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```

# move to  phyloseq section after data selection

# phyloseq section and plot
# color by ecosite (M, N, O, T) and shape by month
```{r}
pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[4], greenpal1[5], greenpal1[6], bluepal[4], bluepal[5], bluepal[6], greenpal2[2])

# taxmat
taxmat<-as.matrix(taxa)
TAX = tax_table(taxmat) #covert to phyoseq object, OTU table

# otu
otumat<-as.matrix(otus_D[,-ncol(otus_D)])
OTU = otu_table(otumat, taxa_are_rows = TRUE)
physeq<-phyloseq(OTU) #TAX removed

sampledata<-sample_data(map_file_D)
mgd<-merge_phyloseq(physeq,sampledata)

mgd.ord <- ordinate(mgd, "NMDS", "bray")
#plot
theme_set(theme_bw())
pmds=plot_ordination(mgd, mgd.ord, color="ecosite", shape="month") #
pmds=pmds + geom_point(size=1.5)+scale_color_manual(values=pal) #
pmds
pmds=pmds +facet_wrap(~ecosite, 5)#
pmds
```

#save plot
```{r}
ggsave(pmds, filename = "16SBacFamily_NMDS_allMonth_by_D1.pdf", width = 8, height = 8, units = "in") #change the depth #1
```

#plot_all samples, with D1-D6
#otus_Ba
```{r}
pal<-c(orangepal[1], orangepal[2], orangepal[3], orangepal[4], orangepal[5], orangepal[6], greenpal1[1], greenpal1[2], greenpal1[3], greenpal1[4], greenpal1[5], greenpal1[6], bluepal[1], bluepal[2], bluepal[3], bluepal[4], bluepal[5], bluepal[6], greenpal2[1], greenpal2[2], greenpal2[3], greenpal2[4], greenpal2[5], greenpal2[6])
# otu
otumat<-as.matrix(otus_Ba[,-ncol(otus_Ba)])
OTU = otu_table(otumat, taxa_are_rows = TRUE)
physeq<-phyloseq(OTU) #TAX removed

sampledata<-sample_data(map_file)
mgd<-merge_phyloseq(physeq,sampledata)

mgd.ord <- ordinate(mgd, "NMDS", "bray")
#plot
theme_set(theme_bw())
pmds=plot_ordination(mgd, mgd.ord, color="eco_dep", shape="depth") #
pmds=pmds + geom_point(size=1.5)+scale_color_manual(values=pal)#
pmds
pmds=pmds +facet_wrap(~ecosite, 5)#
pmds
```

