---
title: "OWC_16S_Heatmap_by_depth"
author: "Liupf"
date: "3/18/2019"
output: html_document
---

#library
```{r}
library(phyloseq)
library(ggplot2)
library(RColorBrewer)

library("factoextra")
library("FactoMineR") # PCA function
library("gplots") #heatmap
library("vegan") # prcomp
library('vegetarian') # normalize.rows function
library(dplyr)
library(tibble)
library(dendsort)
```

```{r}
setwd("~/A_Wrighton_lab/Wetland_project/16SrRNA_ana_Adrienne")
```

# mapping file
map_file_clean_Cores.txt
# ASV file, 
feature_table_w_tax_clean_Cores_ge10.txt
# in bash
sed -e '1s/\./-/g' feature_table_w_tax_clean_Cores_ge10.txt >feature_table_w_tax_clean_Cores_ge10f.txt

# tax file
filtered_ge10_trimmed_tax_tab.txt

# read file
```{r}
map_file<-read.delim("map_file_clean_Cores.txt",header=T,row.names=1)
otus<-read.delim("feature_table_w_tax_clean_Cores_ge10f.txt",header=T,row.names=1, check.names=FALSE) 
taxa<-read.delim("filtered_ge10_trimmed_tax_tab.txt",header = T,row.names = 1) #tax_info, 
```

# add eco_dep column in the mapping file
```{r}
library(tidyr)
map_file$ecosite1=gsub("1|2|3","",map_file$ecosite)
map_file = unite(map_file, eco_dep, c(ecosite1, depth), remove=FALSE)
map_file = map_file[order(map_file$eco_dep), ,drop=FALSE] #M (D1-D6), N, OW, T is the order
```

#color code setup
```{r}
library(RColorBrewer)
bluepal<-brewer.pal(6,"Blues") #open water
orangepal<-brewer.pal(6,"Oranges") # mud
greenpal1<-brewer.pal(6,"Greens") # Typha
greenpal2<-brewer.pal(6,"YlGn") # Nelumbo
#pal<-c(orangepal[4], orangepal[5], orangepal[6], greenpal1[2], greenpal1[3], greenpal1[4], bluepal[4], bluepal[5], bluepal[6], greenpal2[6])
```
# separate all depth, keep all ecosite and month
# data selction, separate Archea and Bacteria
```{r}
otus_Ba <- otus[grep("Bacteria", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ]
otus_Ar <- otus[grep("Archaea", otus$taxonomy, ignore.case=TRUE, invert=FALSE), ] 
#map_file <- map_file[grep("Aug-|Aug_", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
#otus <- otus[, grep("Aug-|Aug_|taxonomy", colnames(otus), ignore.case=TRUE, invert=FALSE)]
```

# convert to relative abundance, not used
```{r}
library(gtools)
#otus_Ba_ReAb=sweep(otus_Ba[,1:(ncol(otus_Ba)-1)], 2, colSums(otus_Ba[,1:(ncol(otus_Ba)-1)]), FUN="/")
#otus_Ba_ReAb$taxonomy= otus_Ba$taxonomy
#otus_Ar_ReAb=sweep(otus_Ar[,1:(ncol(otus_Ar)-1)], 2, colSums(otus_Ar[,1:(ncol(otus_Ar)-1)]), FUN="/")
#otus_Ar_ReAb$taxonomy= otus_Ar$taxonomy
#
#feature_table_Arc_w_tax= merge(otus, Arc, by=0, all=FALSE)
#otus_D_ReAb = otus_D_ReAb [apply(otus_D_ReAb, 1, max)>=0.001,]
#otus_D_ReAb 
#write.table(data.frame("Sample_ID"=rownames(map_file_D),map_file_D),"map_file_Aug_D1_Cores.txt", sep="\t", row.names=FALSE, quote = FALSE)
#write.table(data.frame("OTU_ID"=rownames(otus_D_ReAb ), otus_D_ReAb ),"feature_table_Aug_D1_Cores_ge10.txt", sep="\t", row.names=FALSE, quote = FALSE)
```
# All-month-Bacterial
# D-1
```{r}
#map_file_D <- map_file[grep("-D1-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D1-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```
#move to pca selection and ploting after each data prep of one depth

# D-2
```{r}
#map_file_D <- map_file[grep("-D2-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D2-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```

# D-3
```{r}
#map_file_D <- map_file[grep("-D3-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D3-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```


# D-4
```{r}
#map_file_D <- map_file[grep("-D4-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D4-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```


# D-5
```{r}
#map_file_D <- map_file[grep("-D5-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D5-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```


# D-6
```{r}
#map_file_D <- map_file[grep("-D6-", row.names(map_file), ignore.case=TRUE, invert=FALSE),]
otus_D <- otus_Ba[, grep("-D6-|taxonomy", colnames(otus_Ba), ignore.case=TRUE, invert=FALSE)]
```


###the same to all depth
#loading calculation
```{r}
otus_Dt <- as.data.frame (t(otus_D[,-ncol(otus_D)])) #samples as row, otu as column
otus_Dt.na <- na.omit(otus_Dt)
#hellinger distance
otus_Dt.na.hellinger <- decostand(otus_Dt.na, "hellinger") # calculate hellinger distances
otus_Dt.na.prcomp<-prcomp(otus_Dt.na.hellinger)
aload <- abs(otus_Dt.na.prcomp$rotation)
absloading=sweep(aload, 2, colSums(aload), "/")
absloading = absloading [,1:2]


#the same to all depth
#pca selection
#

#filter top 100 for pc1
absloading_pc1=data.frame(absloading[,1])
absloading_pc1<- absloading_pc1 %>% rownames_to_column('OTUs') %>% top_n (100) %>% column_to_rownames('OTUs')
#filter top 100 for pc2
absloading_pc2=data.frame(absloading[,2])
absloading_pc2<- absloading_pc2 %>% rownames_to_column('OTUs') %>% top_n (100) %>% column_to_rownames('OTUs')
top_otus= unique(c(row.names(absloading_pc1),row.names(absloading_pc2)))

#selected by top_otus names
otus_Dt.na_top=subset(otus_Dt.na,select = top_otus)

#relative abundance cal
Re_abundances<-normalize.rows(otus_Dt.na_top)
percentages<-100*Re_abundances
rownames(percentages)<-rownames(otus_Dt.na_top)
colnames(percentages)<-colnames(otus_Dt.na_top)


#distance for row and col clustering

drows = dist(as.matrix(decostand(percentages, "hellinger")), method = "euclidean") #decostand(OTUtable_t.na, "hellinger")
dcols = dist(t(as.matrix(decostand(percentages, "hellinger"))), method = "euclidean") #decostand(OTUtable_t.na, "hellinger")


#clustering
## prepare for the cluster
##http://slowkow.com/notes/heatmap-tutorial/

#library(dendsort)
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dcols)
mat_cluster_rows <- hclust(drows)
#plot(mat_cluster_cols, main = "", xlab = "", sub = "")



#break setting
# break 2
quantile.range <- quantile(as.matrix(percentages), probs = seq(0, 1, 0.01))
mat_breaks <- seq(quantile.range["5%"], quantile.range["95%"], 0.001)
color  <- colorRampPalette(c("blue", "white","firebrick3"))(length(mat_breaks) - 1); c("royalblue4","#FFFFBF","orangered2")
```

#pheatmap plotting
```{r}
pheatmap(as.matrix(percentages),color = color,breaks= mat_breaks,clustering_distance_rows = drows,clustering_distance_cols=dcols,clustering_method="complete",cellwidth=3,cellheight=5,treeheight_col=0, cluster_cols= mat_cluster_cols,cluster_rows = mat_cluster_rows,show_colnames=FALSE,fontsize=6) # better # hhBC, hclust, hellinger,Bray, complete
#filename="Utah_Arc_OTU_top25_hhBC_reorder.pdf"
```


#retrieve the ordre of OTUs
```{r}
Ad=as.dendrogram(mat_cluster_cols )
order.dendrogram(Ad)
order=data.frame(row.names = colnames(percentages),order.dendrogram(Ad))
order=merge(otus_D,order,by=0)
order_tax=order[ ,c(1,(ncol(order)-1):ncol(order))]
```

#export order data
```{r}
write.table(order_tax, sep="\t",file="BacDep6_otus_hcluster_order_tax.txt",row.names = FALSE) #change the depth1, 2, 3, 4, 5, 6
```